Synthesis of Linear Functional Programs

### 2021-05-06 article/article.tex
tweaks

\usepackage{cmll}
\usepackage{mathpartir}

@@

% General motivation?
Program synthesis is an automated or semi-automated process of
 deriving a program (generating code) from a high-level specification.
 Specifications appear in many forms (e.g.
 polymorphic refinement types~\cite{DBLP:conf/pldi/PolikarpovaKS16},
 examples~\cite{DBLP:conf/popl/FrankleOWZ16},
 graded linear types~\cite{DBLP:conf/lopstr/HughesO20}, ...).
 In this work the synthesis specification will be a linear type, ?possibly with refinements.?
 The chosen specification form dictates the synthesis process, and different ones might
 be approached in completely different ways.
 In spite of the specification, most program synthetizer's have to deal with two main
 inherent complexities: searching over a vast space of valid programs, and interpreting user intent.
 Synthesis has many motivations, it can be seen as a means to improve a programmer's productivity and/or
 program correctness (i.e. through suggestion and/or autocompletion), or as a tool to automate certain parts
 of the programming process (e.g. in the same way AI might generate some boilerplate text to auxiliate a journalist, rather
 than to replace), just to name a few examples.

@@

A type system can be formally described through natural deduction -- as a set of rules that determines the concept of deduction for some language~\cite{prawitznd65}.
The language -- the propositions of a type system are the types themselves, the rules define how to construct types, and being able to deduct a type (i.e. constructing a derivation) proves the type is valid.
[Preciso de ajuda nesta última frase, não é isto que quero dizer mas não sei descrever de forma correta e satisfatória :)]
The natural deduction judgment is extended to include a \emph{term} (i.e. to make proofs explicit with a formal definition), and modified to include a list of hypothesis.
[Também preciso de ajuda a reescrever a última, eu bem que procuro muito mas precisava de alcançar uma mestria de natural deduction - será mais rápido o professor ajudar-me a dar nomes às coisas :)]
This extended judgment, called a typing judgment has the form:
\[
    \Gamma \vdash M : A
\]
Where the natural deduction list of hypothesis, $\Gamma$, is now called \emph{typing environment} (i.e. set of variables and their types), $M$ is still a \emph{term} but also known as ``program'', and $A$, the natural deduction proposition, is now called \emph{type}. This judgment can be read -- $M$ has type $A$ in $\Gamma$.
% As a concrete example, let us describe the typing rule for the tensor type $A \tensor B$

>>>>

\usepackage{cmll}
\usepackage{mathpartir}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{todonotes}

@@

% General motivation?
Program synthesis is an automated or semi-automated process of
deriving a program (i.e.~generating code) from a high-level specification.
Specifications can take many forms (e.g.  polymorphic refinement
types~\cite{DBLP:conf/pldi/PolikarpovaKS16},
examples~\cite{DBLP:conf/popl/FrankleOWZ16}, graded linear
types~\cite{DBLP:conf/lopstr/HughesO20}, ...).  In this work we
explore type-based synthesis where specifications take the form of
linear types.\todo{O que é um linear type?}
%
The chosen specification form dictates the synthesis process, and
different ones might be approached in completely different
ways.\todo{Esta frase não parece ter grande utilidade. O que queres
  dizer aqui?}
%
Regardless of the kind of specification, program synthesizers have to
deal with two main inherent sources of complexity: searching over a
vast space of (potentially) valid programs, and interpreting user
intent.
%
Synthesis has many motivations, it can be seen as a means to
improve a programmer's productivity and/or program correctness
(i.e. through suggestion and/or autocompletion), or as a tool to
automate certain parts of the programming process (e.g. in the same
way AI might generate some boilerplate text to auxiliate a journalist,
rather than to replace), just to name a few examples.
\todo[inline]{Acho que estes 3 ``paragrafos'' deviam estar por ordens diferentes.}

@@

A type system can be formally described through a set of inference
rules that inductively define a judgment of the form $\Gamma \vdash M
: A$, stating that program expression $M$ has type $A$ according to
the \emph{typing assumptions} for variables tracked in $\Gamma$.
\dots

\todo[inline]{Daqui, para tipificar uma função, para Curry-Howard,
  para lógica linear, para proof search, etc\dots Só faz sentido falar de dedução natural no
  contexto de ``lógica'' / Curry-Howard.}

natural deduction -- as a set of rules that determines the concept of deduction for some language~\cite{prawitznd65}.
The language -- the propositions of a type system are the types themselves, the rules define how to construct types, and being able to deduct a type (i.e. constructing a derivation) proves the type is valid.
[Preciso de ajuda nesta última frase, não é isto que quero dizer mas não sei descrever de forma correta e satisfatória :)]
The natural deduction judgment is extended to include a \emph{term} (i.e. to make proofs explicit with a formal definition), and modified to include a list of hypothesis.
[Também preciso de ajuda a reescrever a última, eu bem que procuro muito mas precisava de alcançar uma mestria de natural deduction - será mais rápido o professor ajudar-me a dar nomes às coisas :)]
This extended judgment, called a typing judgment has the form:
\[
    \Gamma \vdash M : A
\]
Where the natural deduction list of hypothesis, $\Gamma$, is now called \emph{typing environment} (i.e. set of variables and their types), $M$ is still a \emph{term} but also known as ``program'', and $A$, the natural deduction proposition, is now called \emph{type}. This judgment can be read -- $M$ has type $A$ in $\Gamma$.
% As a concrete example, let us describe the typing rule for the tensor type $A \tensor B$
